\PYGEOMETRY{}: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-09-19 public/paper/pyg4ometry.tex
Paper update (fixed Stuarts last comments)

There are existing codes that have functionality similar to \PYGEOMETRY{}. ROOT~\cite{Brun:1997pa}, the high-energy physics data analysis framework, 
can load, display and manipulate GDML-like geometry. Tools exist to convert CAD files to GDML, open source examples are GUIMesh~\cite{GUIMesh_Pinto}, commercial 
products include ESABASE2~\cite{ESABASE2} and FASTRAD~\cite{FASTRAD}. There are also tools from the fusion and neutronics community that can convert CAD 
geometry into formats usable by MCRT codes, with examples including DAGMC~\cite{DAGMC} and McCAD~\cite{McCad}. In principle CAD software can export shape data to STL 
(or other similar mesh formats), which can be used by Geant4~\cite{poole2012acad}. In practice, using a lot of CAD models is difficult if that model is comprised of a 
large number of parts. Exporting, assigning material to, and placing the STL components into the MCRT code can be a cumbersome task. The existent set of software does not 
provide a complete set of tools to efficiently create complex geometries.

@@

There are many output format extensions that can be considered for \PYGEOMETRY{}. 
Geant4 geometry is primarily created by writing C++ programmes, so an output writer that 
converts the \PYGEOMETRY{} in-memory representation to C++ will allow rapid geometry 
modelling but . This is not implemented in the current version but could be relatively quickly implemented
for users that require this functionality. At present \PYGEOMETRY{} supports reading and writing 
Geant4 (GDML) and FLUKA files but could be extended without significant effort to other MCRT codes 
like MCNP.

@@

The conversion which would most dramatically enhance the creation of geometry is CAD to Geant4 or FLUKA
without use use of a triangular or tetrahedral mesh. There are existing approaches to decompose BREP solids to Geant4 and 
FLUKA-like CSG geometry~\cite{WangNuclSciTech31-82-2020, LuFusionEngineeringAndDesign124-2017}. 
The FreeCAD/OpenCASCADE interface combined with the Geant4 and FLUKA Python API in \PYGEOMETRY{} 
will allow for the creation of CAD BREP decomposition algorithms.

@@

The development of \PYGEOMETRY{} has received funding from Science and
Technology Research council grant ``The John Adams Institute for
Accelerator Science'' ST/P00203X/1 through the John Adams Institute at
Royal Holloway, and Royal Holloway Impact Acceleration Account.  Multiple
undergraduate and masters students have contributed to the code for their
degree project work: Simon Williams, Benjamin Shellswell and Joshua Albrecht.
STB has benefited from discussions with and insight of Vasilis Vlachoudis
(CERN) on FLUKA geometry and Flair.  Cédric Hernalsteens (ULB/CERN) and
Alistair Butcher (RHUL) also made useful contributions to the code.

>>>>

There are existing codes that have functionality similar to \PYGEOMETRY{}. ROOT~\cite{Brun:1997pa}, the high-energy physics data analysis framework, 
can load, display and manipulate GDML-like geometry. Tools exist to convert CAD files to GDML, open source examples are GUIMesh~\cite{GUIMesh_Pinto}, commercial 
products include ESABASE2~\cite{ESABASE2} and FASTRAD~\cite{FASTRAD}. There are also tools from the fusion and neutronics community that can convert CAD 
geometry into formats usable by MCRT codes, with examples including DAGMC~\cite{DAGMC} and McCAD~\cite{McCad}. In principle CAD software can export shape data to STL 
(or other similar mesh formats), which can be used by Geant4~\cite{poole2012acad}. In practice, using a lot of CAD models is difficult if that model is comprised of a 
large number of parts. Exporting, assigning material to, and placing the STL components into the MCRT code can be a cumbersome task. Almost all modern CAD tools like 
CATIA, Inventor, SolidWorks have a scripting language to allow users to programatically generate geometry. This scripting functionality is not available with either 
GDML or FLUKA were users write the geometry files directly. The existent set of software does not provide a complete set of tools to efficiently create complex geometries.

@@

There are many output format extensions that can be considered for \PYGEOMETRY{}. 
Geant4 geometry is primarily created by writing C++ programmes, so an output writer that 
converts the \PYGEOMETRY{} in-memory representation to C++ will allow rapid geometry 
modelling but inclusion of the geometry into an existing Geant4 application. This is not 
implemented in the current version but could be relatively quickly implemented for users that 
require this functionality. At present \PYGEOMETRY{} supports reading and writing 
Geant4 (GDML) and FLUKA files but could be extended without significant effort to other MCRT codes 
like MCNP.

@@

The conversion which would most dramatically enhance the creation of RTMC geometry is CAD to Geant4 or FLUKA
without use use of a triangular or tetrahedral mesh. There are existing approaches to decompose BREP solids to Geant4 and 
FLUKA-like CSG geometry~\cite{WangNuclSciTech31-82-2020, LuFusionEngineeringAndDesign124-2017}. 
The FreeCAD/OpenCASCADE interface combined with the Geant4 and FLUKA Python API in \PYGEOMETRY{} 
will allow for the creation of CAD BREP decomposition algorithms. There are python based CAD modelling tools like 
cadquery \cite{cadquery} which allow the creation of models using pure Python which should allow the conversion of GDML to STEP.

@@

The development of \PYGEOMETRY{} has received funding from Science and
Technology Research council grant ``The John Adams Institute for
Accelerator Science'' ST/P00203X/1 through the John Adams Institute at
Royal Holloway, and Royal Holloway Impact Acceleration Account. Gian Luigi D'Alessandro (CERN/JAI)
provided geometry and tested multiple conversions of beamlines. Multiple
undergraduate and masters students have contributed to the code for their
degree project work: Simon Williams, Benjamin Shellswell and Joshua Albrecht.
STB has benefited from discussions with and insight of Vasilis Vlachoudis
(CERN) on FLUKA geometry and Flair.  Cédric Hernalsteens (ULB/CERN) and
Alistair Butcher (RHUL) also made useful contributions to the code.
